---
title: "Rowery Waszyngton"
author: "Michał Syrkiewicz"
output: pdf_document
---
Loading used libraries. Devtools used to load "R-package" my own small package made for this assignment

```{r}
library(dplyr)
library(stringr)
library(devtools)
load_all("R-package")
```
```{r}
help(wyodrebnij)
help(czas)
help(unormuj)
help(correlation)
```

Reading and saving quarterly data from Washington bikes
All needed csv-s in zip folder.
```{r}
Q1 <- read.csv("2013Q1-capitalbikeshare-tripdata.csv")
Q2 <- read.csv("2013Q2-capitalbikeshare-tripdata.csv")
Q3 <- read.csv("2013Q3-capitalbikeshare-tripdata.csv")
Q4 <-read.csv("2013Q4-capitalbikeshare-tripdata.csv")
```
Extracting month and year from every dataset to group by day.
Making two separate frames for "Members" and "Casual" users.

```{r}
Q1 <- wyodrebnij(Q1,name="Start.date",from="%Y-%m-%d",what = c("%d","%m"), newName = c("Day","Month"))
Q1 <- Q1 %>% group_by(Day,Month) %>% count(Member.type) %>% mutate(Data = paste("2013",Month,Day,sep="-"))
casual <- Q1[Q1$Member.type=="Casual",]
registere <<- Q1[Q1$Member.type=="Member",]
```

```{r}
Q2 <- wyodrebnij(Q2,name="Start.date",from="%Y-%m-%d",what = c("%d","%m"), newName = c("Day","Month"))
Q2 <- Q2 %>% group_by(Day,Month) %>% count(Member.type) %>% mutate(Data = paste("2013",Month,Day,sep="-"))
casual1 <- Q2[Q2$Member.type=="Casual",]
registere1 <- Q2[Q2$Member.type=="Member",]
```

```{r}
Q3 <- wyodrebnij(Q3,name="Start.date",from="%Y-%m-%d",what = c("%d","%m"), newName = c("Day","Month"))
Q3 <- Q3 %>% group_by(Day,Month) %>% count(Member.type)  %>% mutate(Data = paste("2013",Month,Day,sep="-"))
casual2 <- Q3[Q3$Member.type=="Casual",]
registere2 <- Q3[Q3$Member.type=="Member",]
```

```{r}
Q4 <- wyodrebnij(Q4,name="Start.date",from="%Y-%m-%d",what = c("%d","%m"), newName = c("Day","Month"))
Q4 <- Q4 %>% group_by(Day,Month) %>% count(Member.type) %>% mutate(Data = paste("2013",Month,Day,sep="-"))
casual3 <- Q4[Q4$Member.type=="Casual",]
registere3 <- Q4[Q4$Member.type=="Member",]
```

Merging all data from year 2013

```{r}
casuals<-rbind(casual,casual1,casual2,casual3)
registered<-rbind(registere,registere1,registere2,registere3)
```
Reading and saving weather data from Washington 2013
```{r}
pogoda2013 <- read.csv("2013_DC.csv",skip=3)
```
Small aestethic changes and joining weather and bike data
```{r}
colnames(registered)[4]<- "registered"
colnames(casuals)[4]<- "casual"
registered<- registered[,c(-3)]
casuals<- casuals[,c(-3)]
wynik<- inner_join(casuals, pogoda2013, join_by(Data == time))
wynik <- inner_join(wynik,registered, by=(join_by(Day,Month,Data)))
```
Looking for correlation in data

```{r}
wynik <- wynik[,c(-1,-2)]
wynik<- wynik[,c(2,1,20,3:19)]
correlation(wynik, c(2:3), c(4:20), high=0.65)
```
As you can see we have high correlation with evapotranspiration, temperature and shortwave radiation. I'll try making new columns with these columns
```{r}
wynik <- wynik %>% mutate(Stosunki1 = temperature_2m_max...C./et0_fao_evapotranspiration..mm., Stosunki2 = temperature_2m_mean...C./et0_fao_evapotranspiration..mm., Stosunki3 = apparent_temperature_max...C./et0_fao_evapotranspiration..mm.,
                          Stosunki4 = shortwave_radiation_sum..MJ.m../et0_fao_evapotranspiration..mm., Inne = shortwave_radiation_sum..MJ.m..*apparent_temperature_max...C.,
                          Inne2 = shortwave_radiation_sum..MJ.m..*temperature_2m_max...C., Inne3 = shortwave_radiation_sum..MJ.m..*et0_fao_evapotranspiration..mm.)
correlation(wynik,c(2,3),c(4:27),high=0.7)
```
As you can see we have higher correlation with products of these columns trying one more time.
```{r}
wynik <- wynik %>% mutate(Inne4 = Inne2 * et0_fao_evapotranspiration..mm.)
correlation(wynik,c(2,3),c(4:28),high=0.7)
```
Let's plot those columns in order to find more correlation inside data.

```{r}
par(mfrow = c(2, 3))
plot(wynik$casual,wynik$et0_fao_evapotranspiration..mm., xlab = "Casuals", ylab="Evapotranspiration")
plot(wynik$casual,wynik$temperature_2m_max...C., xlab = "Casuals", ylab="Temperature")
plot(wynik$casual,wynik$shortwave_radiation_sum..MJ.m.., xlab = "Casuals", ylab="Radiation")
plot(wynik$casual,wynik$Inne, xlab = "Casuals", ylab="Apparent Temperature * Radiation")
plot(wynik$casual,wynik$Inne2, xlab = "Casuals", ylab="Temperature * Radiation")
plot(wynik$casual,wynik$Inne4, xlab = "Casuals",ylab="TRE")
```
As we can see days with more than around 2400 casual bikers tend to be different than most day of the year. Let's eliminate them.
```{r}
wynik1<-wynik[wynik$casual<2400,]
correlation(wynik1,c(2,3),c(4:28),high=0.75, method="spearman")

```
Let's normalize data in second data frame
```{r}
wynik2 <- unormuj(wynik1,which = c(2:28),check=FALSE,decimals = 7)
par(mfrow = c(2, 2))
plot(wynik2$casual,wynik2$Inne4, xlab = "Casuals",ylab="TRE")
plot(wynik2$casual,wynik2$temperature_2m_max...C., xlab = "Casuals", ylab="Temperature")
plot(wynik2$casual,wynik2$Inne2, xlab = "Casuals", ylab="Temperature * Radiation")
plot(wynik2$casual,wynik2$Inne, xlab = "Casuals",ylab="Apparent Temperature * Radiation")
correlation(wynik1,c(2,3),c(4:28),high=0.75, method="spearman")
```
Let's standarize data in 3 data frame
```{r}
wynik3<-as.data.frame(sapply(wynik1[,-1], function(data) (data-mean(data))/sd(data)))
par(mfrow = c(2, 2))
plot(wynik3$casual,wynik3$Inne4, xlab = "Casuals",ylab="TRE")
plot(wynik3$casual,wynik3$temperature_2m_max...C., xlab = "Casuals", ylab="Temperature")
plot(wynik3$casual,wynik3$Inne2, xlab = "Casuals", ylab="Temperature * Radiation")
plot(wynik3$casual,wynik3$Inne, xlab = "Casuals",ylab="Apparent Temperature * Radiation")
correlation(wynik1,c(2,3),c(4:28),high=0.75, method="spearman")

```

